Collective behavior of Janus Particles Suspended in a Viscous Fluid

### 2022-06-10 doc/PRF2022/PRF_main.tex
added rewrite of governing eqs to top of PRF_main.tex

{\bf More results:}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{multilamellar_shear_Energy}
    \includegraphics[width=0.4\textwidth]{multilamellar_shear_MSD}
  \end{center}
      \caption{A multi-lamellar structure in shear flows with shear rates $\dot\gamma=\{0.05, 0.1, 0.15\}$. We compute the energy and the mean square displacement $MSD(t) = \frac{1}{N_b^2}\sum_{i=1}^{N_b} ||D(t)-D(0) ||_2$ where the matrix $D$ is the pairwise distance matrix at time $t$.    }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{multilamellar_shear0p1_Energy_LsFs}
    \includegraphics[width=0.4\textwidth]{multilamellar_shear0p1_MSD_LsFs}
  \end{center}
    \caption{ A multi-lamellar structure in a shear flow when the shear rate is $\dot\gamma=0.1$. We tune the length scale parameters, radius, $\rho$ and $\rho_0$, by $\pm 5\%$ and the force scale parameters, $\gamma$ and $M$, by $\pm 5\%$.    }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{multilamellar_TG_Energy}
    \includegraphics[width=0.4\textwidth]{multilamellar_TG_MSD}
  \end{center}
    \caption{A multi-lamellar structure in a Taylor-Green flow when the flow strength $V_0=\{0.1,0.15,0.2\}$. }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{checker_board_shear_Energy}
    \includegraphics[width=0.4\textwidth]{checker_board_shear_MSD}\\
  \includegraphics[width=0.4\textwidth]{checker_board_shear0p125_OP}
    \includegraphics[width=0.4\textwidth]{checker_board_shear0p15_OP}
  \end{center}
\caption{A checker board in shear flows with shear rates $\dot\gamma=\{0.125, 0.15\}$. We propose the order parameter defined by $S_2 = \frac12(3\cos(\theta-\bar\theta)^2-1)$}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[width=0.4\textwidth]{checker_board_TG_Energy}
    \includegraphics[width=0.4\textwidth]{checker_board_TG_MSD}\\
      \includegraphics[width=0.4\textwidth]{checker_board_TG0p2_OP}
    \includegraphics[width=0.4\textwidth]{checker_board_TG0p3_OP}
  \end{center}
\caption{A checker board in Taylor-Green flows when the flow strength $V_0=\{0.2,0.3\}$. }
\end{figure}

>>>>

\section{Governing Equations: Hydrophobic Attraction Potential Mobility Problem} 
The governing equations are a nonlinear system
for the position and orientation of a collection of
rigid Janus particles.
We first pose the Stokes equations
for the mobility problem giving the hydrodynamic interactions
for the particle suspension.  
The hydrophobic forces come from solving a screened Laplace equation.
Particle collisions are avoided through a near-field, pair
potential.

\subsection{Mobility problem}
A suspension of $N_b$ Janus particles is immersed in a vicsous solvent.
The domain $\Omega(t) \subset \mathbb{R}^2$ denotes the solvent phase
where $t$ is time.
The particles are disks of radius $c$ with center $\aa_i$ and
orientation $\theta_i$ relative to a fixed horizontal axis.  
The boundary of the $\Omega$ is $\bd\Omega = \Gamma_1 \cup
\cdots \cup \Gamma_{N_b}$, where $\Gamma_i$ is the boundary of Janus
particle $i$. Assuming inertial terms are
negligible, we have 
\begin{alignat}{3}
\label{eq:stokes}
  -\mu \Delta \uu + \nabla p &= \mathbf{0},     && \xx \in \Omega, \\
  \nabla\cdot \uu &= 0, \qquad && \xx \in \Omega, \\
  \uu - \uu_\infty &\to \mathbf{0}, && |\xx| \to \infty,
\end{alignat}
where $\uu$ is the velocity of the solvent,
$p$ is the pressure, $\uu_\infty$ is the
background flow, and $\mu$ is the constant viscosity.
On the boundaries, the solvent
velocity satisfies a no-slip boundary condition for a rigid body motion
\begin{equation}
  \label{eq:rigid-vel}
  \vv_i + \omega_i (\xx - \aa_i)^\perp = \uu(\xx), 
    \quad \xx \in \Gamma_i,
\end{equation}
where $\vv_i$ is the translational velocity, $\omega_i$ is the
angular velocity, and
$\langle x, y \rangle^{\perp} = \langle -y, x\rangle$.

To close the system \eqref{eq:stokes}-\eqref{eq:rigid-vel},
we use
\begin{equation}
  \label{eq:force}
  \int_{\Gamma_i} \ssigma \cdot \nnu \, \dif s = \FF_i, \quad
  \int_{\Gamma_i} (\xx - \aa_i)^\perp \cdot 
  (\ssigma \cdot \nnu) \, \dif s = T_i,
  \quad i=1,\ldots,N_b,
\end{equation}
where $\FF_i \in \mathbb{R}^2$ and $T_i \in \mathbb{R}$
are interparticle forces and torques. 
Here $\ssigma = -p \mathbf{I} + \mu \left(\nabla \uu + \nabla \uu^T
\right)$ is the hydrodynamic stress tensor (pressure tensor) and
$\nnu$ is the particle outward normal.

To get $\FF_i$ and $T_i$,
the particles minimizes the free energy
\begin{align}
\label{eq:free_energy}
F[u] = \gamma
\int_{\Omega} \rho |\nabla u|^2 + \rho^{-1} u^2
\,d\xx
+ \frac{M}{2}
\sum_{j \neq i} 
P\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right), \quad
\end{align}
Here, $u(\xx,t)$ is an order parameter for the structure of water
in the presence of hydrophobic surfaces,
$\rho$ is a decay length, and $\gamma$ is an interfacial tension. 
We model water structure
using the screened Laplace equation boundary value problem:
\begin{gather}
  \label{eq:SL}
  -\rho^2 \Delta u + u = 0,\quad \xx \in \Omega, \\
  u = g, \quad \xx \in \bd\Omega, \quad 
  u \rightarrow 0, \quad |\xx| \rightarrow \infty.
\end{gather}
The boundary condition $g$ represents the degree of hydrophobicity
the particle (Figure~\ref{fig:bcs}).

Variation of the
the free energy \eqref{eq:free_energy}
provides
\begin{equation}
  \FF_i = \int_{\Gamma_i} \mathbf{T}\nnu \, \dif s
  - \frac{M}{\rho_0}
  \sum_{j \neq i}
  \frac{\aa_i - \aa_j}{|\aa_i - \aa_j|}
P'\left(\frac{|\aa_i - \aa_j|-2c}{\rho_0}\right)
  ,\quad
T_i = \int_{\Gamma_i} (\xx - \aa_i)^{\perp}\cdot (\mathbf{T}\nnu) \, \dif s, 
\end{equation}
for $i = 1,\dots,N_b$. The term 
$
\mathbf{T}
= \gamma \left[ \rho^{-1} u^2 \mathbf{I}
  + \rho \left(|\nabla u|^2 \mathbf{I} - 2\nabla u \nabla u^T\right)\right]
$
is the hydrophobic stress it is derived from 
variation of the domain as described in ~\cite{Fu2018_SIAM}.

The dimensionless repulsion profile $P$
takes the form
$P(s) = 1 - \sin(\pi s/2)$ for $0 \leq s < 1$
and $P(s) = 0$ for $1 \leq s < \infty$.
As such, only those particles within
a $\rho_0$-tubular neighborhood of $\Gamma_i$
contribute to repulsion.  The parameter $M$ is the repuslion modulus.

\subsection{Boundary integral represenation and time stepping}
We recast \eqref{eq:stokes}-\eqref{eq:rigid-vel} and \eqref{eq:SL}
as boundary integral equations (BIEs) and discretize each BIE at
$N$ points on each of the $N_b$ particles with a collocation method.
To express the solution of \eqref{eq:SL},
we adopt the double layer potential
\begin{align}
\label{eq:SL_BIE}
u({\bf x}) = \DD[\sigma](\xx) = \int_\Gamma \frac{\partial G(\xx-\yy)}{\partial \nnu_\yy}\sigma(\yy)\, \dif s_\yy
, \quad \xx \in \Omega,
\end{align}
where $G(\xx)$ is the fundamental solution to the screened Laplace equation \eqref{eq:SL}
and where $\sigma$ is a scalar-valued density function. 
For the velocity, we use the representation  
\begin{align}
  \label{eq:velocity}
  \uu(\xx) = \uu_\infty(\xx) + \DD[\eeta](\xx) + 
    \sum_{i=1}^{N_b} \left(\SS(\xx,\aa_i) \cdot \FF_i + 
    \RR(\xx,\aa_i) T_i\right), \quad \xx \in \Omega,
\end{align}
with the vector-valued density function $\eeta$,
Stokeslet 
$\SS(\xx,\aa_i) = \frac{1}{4\pi} \left(-\log |\rr|\mathbf{I} + \frac{\rr \otimes \rr}{|\rr|^2}\right)$,
and Rotlet
$\RR(\xx,\aa_i) = \frac{1}{4\pi} \frac{\rr^\perp}{|\rr|^2}$.

Finally, together with \eqref{eq:velocity},
\eqref{eq:stokes}-\eqref{eq:rigid-vel} become 
\begin{alignat}{3}
  \label{eq:vel_BLM_rep}
  \vv_i + \omega_i (\xx - \aa_i)^\perp &= \uu_\infty(\xx)
    -\frac{1}{2} \eeta(\xx) + \DD[\eeta](\xx) 
    + \sum_{j=1}^{N_b} 
    \left(\SS(\xx,\aa_j) \cdot \FF_j + \RR(\xx,\aa_j) T_j\right),\\
  \label{eq:mobility}
  \int_{\Gamma_i} \eeta \, \dif s &= \mathbf{F}_i, \quad
  \int_{\Gamma_i} \eeta \cdot (\xx-\aa_i)^\perp \, \dif s = T_i,
\end{alignat}
for $\xx \in \Gamma_i$ and $i = 1,\dots,N_b$.

We discretize \eqref{eq:SL_BIE} and \eqref{eq:vel_BLM_rep}
using high-order interpolation-based quadrature rules.
Integrals that are smooth are computed with
the spectrally-accurate trapezoid rule, and nearly-singular integrals,
caused by close contact between two particles, are computed with a
high-order interpolation-based quadrature rule (\cite{qua-bir2014}).

After discretizing and applying quadrature,
the result is an $NN_b \times NN_b$ and $2NN_b \times 2NN_b$
linear system for \eqref{eq:SL_BIE} and \eqref{eq:vel_BLM_rep}, respectively.
These are solved with matrix-free GMRES, and we guarantee that the number of
GMRES iterations is mesh-independent by using second-kind BIEs.

Once determined the translational and angular velocities $\vv_i, \omega_i$
have been determined, a second-order Adams-Bashforth scheme updates the particle
positions and orientations:
\begin{equation}
  \label{eq:adams-bashforth}
  \begin{aligned}
    \aa_i(t_{n+2}) &=
\aa_i(t_{n+1})
+ \tfrac{3}{2}\Delta t \vv_i(t_{n+1})
- \tfrac{1}{2}\Delta t \vv_i(t_{n}),\\
    \theta_i(t_{n+2}) &=
\theta_i(t_{n+1})
+ \tfrac{3}{2}\Delta t \omega_i(t_{n+1})
- \tfrac{1}{2}\Delta t \omega_i(t_{n}).
  \end{aligned}
\end{equation}

\begin{figure}
\begin{center}
  \includegraphics[width=0.24\textwidth]{Fig2a.png}
  \includegraphics[width=0.24\textwidth]{Fig2b.png}
  \includegraphics[width=0.24\textwidth]{Fig2c.png}
  \includegraphics[width=0.24\textwidth]{Fig2d.png}
\end{center}
\begin{caption}{\label{fig:relax}
  Panel (a) shows 60 particles confined to a square with random 
  positions and orientation.
  In a quiescent background flow, 
  the self-assembled particles take on distinct configurations
  depending on the value of the shift parameter $b$.}
\end{caption}
\end{figure}

\section{Equilibrium configurations}
\subsection{Boundary conditions and normalization}
The present paper considers boundary condition of the form
\begin{equation}
  \label{eq:bc-type}
g(\xx) = a(b + \cos \alpha),\quad a = (\pi c(2b^2 + 1))^{-1/2},\quad \xx \in \Gamma_i,
\end{equation}
where $\alpha$ is the angle between $\xx - \aa_i$ and the particle director
$\dd_i = (\cos \theta_i, \sin \theta_i)$.
We refer to the side of the particle where $\alpha = 0$ as the head and the
side where $\alpha = \pi$ as the tail. 
The boundary conditions are
normalized so that $\int_{\Gamma_i} g^2(\xx) \, \dif s = 1$.

To see the effect of the shift parameter $b$, we simulate the dynamics of
60 particles in a quiescent flow as shown in Figure~\ref{fig:relax}.
There are three cases:
(i) $b = 1$, bilayers,
(ii) $b = 2$, multilamellar
and (iii) $b = 0$, striated.

Case (i), $b = 1$, mimicks an amphiphilic particle.
The boundary condition $g$ is everywhere nonnegative,
takes the value $0$ on the
hydrophilic tail,
and takes the maximum value $a$ on the
hydrophobic head. 
The interaction is attractive, and particles collectively orient their
hydrophobic sides to form bilayers  (Figure~\ref{fig:relax}(b)).

In case (ii), $b = 1$ the boundary condition is positive everywhere.
It takes the maximum value $3a$ on head where $\alpha = 0$ and the
minimum value $a$ on the tail side where $\alpha = \pi$. This way,
both sides of the Janus particle are hydrophobic but
moreso on the $\alpha = 0$ side.  
Over a short time scale, these particles also self-assemble into bilayers. But unlike for case (i),
the heads are also hydrophobic and so in the long-time dynamics
these bilayers also adhere to one another 
forming multilamella structures. Figure~\ref{fig:relax}(c), for example, shows a
multilamellar structure with 2 layers. The number of folds depends on several factors,
for instances, number of particles and intial configurations.

A water structure with
positive/negative charge is possible in case (iii), $b = 0$
\cite{MaRa76, Ma77}.
Here, the head repels the tail of other particles
and the particles initially form chains with their directors perpendicular
to the length of the chain.  
It is worth noting that in the dipolar molecules modeled by
Laplace equation, the equilibrium configurations consist of strands,
rather than a patch, of particles and also have
alternating orientations \cite{}. Unlike the electrical case, though,
the chains form striated layers
where the particles lie on a 
grid and the orientations alternate between layers (Figure~\ref{fig:relax}(d)).

\section{DRAFT RESULTS BELOW}
{\bf More results:}

\begin{figure}
  \begin{center}
  \includegraphics[height=0.2\textheight]{multilamellar_shear_Energy}
    \includegraphics[height=0.2\textheight]{multilamellar_shear_MSD}\\
          \includegraphics[height=0.2\textheight]{multilamellar_shear_LocPar}
  \end{center}
      \caption{A multi-lamellar structure in shear flows with shear rates $\dot\gamma=\{0.05, 0.1, 0.15\}$. We compute the energy and the mean square displacement $MSD(t) = \frac{1}{N_b^2}\sum_{i=1}^{N_b} ||D(t)-D(0) ||_2$ where the matrix $D$ is the pairwise distance matrix at time $t$.    }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[height=0.2\textheight]{multilamellar_shear0p1_Energy_LsFs}
    \includegraphics[height=0.2\textheight]{multilamellar_shear0p1_MSD_LsFs}
  \end{center}
    \caption{ A multi-lamellar structure in a shear flow when the shear rate is $\dot\gamma=0.1$. We tune the length scale parameters, radius, $\rho$ and $\rho_0$, by $\pm 5\%$ and the force scale parameters, $\gamma$ and $M$, by $\pm 5\%$.    }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[height=0.2\textheight]{multilamellar_TG_Energy}
    \includegraphics[height=0.2\textheight]{multilamellar_TG_MSD}\\
    \includegraphics[height=0.2\textheight]{multilamellar_TG_LocPar}
  \end{center}
    \caption{A multi-lamellar structure in a Taylor-Green flow when the flow strength $V_0=\{0.1,0.15,0.2\}$. }
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[height=0.2\textheight]{checker_board_shear_Energy}
    \includegraphics[height=0.2\textheight]{checker_board_shear_MSD}\\
  \includegraphics[height=0.2\textheight]{checker_board_shear_LocPar}
  \end{center}
\caption{A checker board in shear flows with shear rates $\dot\gamma=\{0.125, 0.15\}$. We propose the order parameter defined by $S_2 = \frac12(3\cos(\theta-\bar\theta)^2-1)$}
\end{figure}

\begin{figure}
  \begin{center}
  \includegraphics[height=0.2\textheight]{checker_board_TG_Energy}
    \includegraphics[height=0.2\textheight]{checker_board_TG_MSD}\\
      \includegraphics[height=0.2\textheight]{checker_board_TG_LocPar}
  \end{center}
\caption{A checker board in Taylor-Green flows when the flow strength $V_0=\{0.2,0.3\}$. }
\end{figure}
